% Documentation of the StartWAM sensor

\subsubsection{\texttt{Sensor.StartWAM}}

\paragraph{Description}

The goal of the StartWAM sensor is to detect the translation start
codons and to give them a score reflecting the context accordance with
given models. A score is attributed at each potential start codons
(ATG), according to Weight Array Method (see Zhang and Marr, {\em
  Comput Appl Biosci.} 1993 Oct;9(5):499-509), or Weighted Array
Matrix models (Salzberg, {\em Comput Appl Biosci} 1997
Aug;13(4):365-76). A WAM describes a consensus motif of a functional
signal, and is composed by one markovian model per each position of
the motif. Here the motif is defined by the ATG (present in all start
codons) plus the two flanking contexts (used by the WAM).  Globally,
the score of a motif is a function of the emission probabilities of
this motif given a true positive model and a false positive model.

The sensor is activated by setting the parameter
\texttt{Sensor.StartWAM.use} to 1.  The user have to specify in
the parameter file the base name (prefix) of the model files
(\texttt{StartWAM.modelfilename}), the size of the context
(\texttt{StartWAM.NbNtBeforeATG, StartWAM.NbNtAfterATG}), the order of
the markovian models \texttt{StartWAM.MarkovianOrder} (the same for
each position of the motif), and the scaling parameters
\texttt{StartWAM.ScaleCoef*, StartWAM.ScalePenalty*}.

Here is an example of StartWAM parameters definition.
\begin{Verbatim}[fontsize=\small]
StartWAM.modelfilename     WAM/WAM.ARA.START  # base name of the model files
StartWAM.NbNtBeforeATG     3                  # amount context
StartWAM.NbNtAfterATG      3                  # aval context
StartWAM.MarkovianOrder    1                  # order of the markovian models
StartWAM.ScaleCoef*        0.1594             # scaling parameter
StartWAM.ScalePenalty*     -3.1439            # scaling parameter
Sensor.StartWAM.use        1               # Use StartWAM sensor
Sensor.StartWAM            1                  # Sensor priority
\end{Verbatim}

\paragraph{Input files format}

This StartWAM sensor requires a true positive and a false positive
model file per motif position. These files have to be present in the
path given by \texttt{StartWAM.modelfilename} from the plugins
directory (see \texttt{EuGene.PluginsDir} parameter). These models can
be generated using \texttt{WAMbuilder.cc} (see
\texttt{eugene/src/SensorPlugins/0\_SensorTk/GetData/README}).  The
file name of a model is a concatenation of the base name (prefix)
specified in the parameter file, an extension (suffix) specified in
the \texttt{WAM.h} file (\texttt{.TP.} for true positive and
\texttt{.FP.}  for false positive), and a number between 00 and 99
indexing the position in the motif (restricting thus the motif length
to a maximum of 100 nt).

As an example, with the base name \texttt{WAM.ARA.START9} (refering to
{\em A.thaliana} models of 9nt-length start motif), one can found these files:
\begin{Verbatim}[fontsize=\small]
WAM.ARA.START9.FP.00
WAM.ARA.START9.FP.01
...
WAM.ARA.START9.FP.07
WAM.ARA.START9.FP.08
WAM.ARA.START9.TP.00
WAM.ARA.START9.TP.01
...
WAM.ARA.START9.TP.08
\end{Verbatim}

These files are in binary form, each containing the properties of a
markovian model (see documentations of \texttt{WAMbuilder.cc} and
\texttt{markov.cc}).

\paragraph{Filtering input information}

Each binary model file is verified when loaded, checking if 3 expected
properties of its markovian model are verified: the order, the
alphabet size, and the total number of possible words (these 3 values
are automatically included during the models generation by
\texttt{WAMbuilder.cc}). This test is done in the loading file method
``\texttt{chargefichier}'' in markov.cc.

\paragraph{Integration of information}

At each ATG of the genomic sequence a score is assigned depending on
the ATG flanking context. If there isn't enough context, e.g. in the
sequence extremities, nothing is done.  This score is provided by a
scaled sum of likelihood ratio, computed as following.

Let be $P^t_i$ the emission probability of the nucleotid at position
$i$ in the motif according to the True Positive model, and $P^f_i$ the
emission probability of the nucleotid given by the False Positive
model. The score given by the WAM for the entire motif $M$ of length
$L$ is:
\[ S_M = \sum_{i=0}^{L} log\left(\frac{P^t_i}{P^f_i}\right) \]

This score is then scaled with the \texttt{StartWAM.ScaleCoef*} and the
\texttt{StartWAM.ScalePenalty*} parameters, following this formula :
$$S_M . \texttt{StartWAMScaleCoef*} + \texttt{StartWAMScalePenalty*}$$

This rescaled score is finally integrated into the \EuGene\ graph on
the UTR5 $\to$ EXON transition edge just before the considered ATG. The
score applies only to the edge corresponding to the situation where
the signal is used. The edge corresponding to the situation where the
signal is not used is unchanged.

\paragraph{Post analyse}

No Post-Analyse.

\paragraph{Graph}

Vertical blue lines (whose length is function of the score) are
plotted on the corresponding frame for each start codon which score is
higher than a defined treshold. This threshold is defined in the
\texttt{StartWAM.cc} file as -\texttt{PlotScoreIncrease}.
