% Documentation of the BlastX sensor

\subsubsection{\texttt{Sensor.BlastX}}

\paragraph{Description}

The BlastX sensor allows to exploit similarities with homologous
proteins. The similarities influence exon and intron detection.
Similarities from several databases can be exploited. Usually 3
databases are used: SwissProt, PIR and TrEMBL.

A label $i$ (that could vary from 0 to 9) is assigned at each
considered database. Files describing a collection of similarities with
a sequence have an extension .blast$<$i$>$ (\texttt{.blast0,...,.blast9}).

The user has to specify the list of labels to consider, the
confidence accorded to each, the minimum length of an intron and a
number of amino acids involved in intron incitation.  The sensor is
activated by either:
\begin{itemize}
\item the \texttt{-b} argument \index{CmdFlags}{[BlastX activation] b}
  that allows to specify the labels to consider, for example
  \texttt{-b092} to use the levels 0, 9, 2 (files \texttt{.blast0,
    .blast9, .blast2}),
\item the value 1 for the parameter \texttt{Sensor.BlastX.use} in
  the parameter file and the labels to consider in the
  \texttt{.BlastX.levels} parameter.
\end{itemize}
The confidence in analyzes have to be specified in the parameter file
giving values to the parameters \texttt{BlastX.level}$<$i$>$*.  The
minimum length of an intron is defined in the \texttt{BlastX.minIn}
parameter.  A number of amino acids defined in the
\texttt{BlastX.blastxM*} parameter that allows to define if 2
similarities are near (see the paragraph Integration of information).
Finally the \texttt{BlastX.postProcess} parameter (when set to 1)
allows to request to analyse how BlastX information are integrated in
the final prediction.

Here is an example of BlastX parameters definition.
\begin{Verbatim}[fontsize=\small]
BlastX.postProcess 1     # analyse prediction accorded to BlastX information
BlastX.levels      012      # use levels 0, 1, and 2
BlastX.activegaps      0    # make gap active on level 0
BlastX.level0*     0.2      # confidence in the level 0
BlastX.level1*     0.0      # confidence in the level 1
BlastX.level2*     0.0      # confidence in the level 2
BlastX.blastxM*    10       # nb of amino acids implicated in intron incitation
BlastX.minIn       50       # minimum length of intron
Sensor.BlastX.use  1     # Use BlastX sensor
Sensor.BlastX      1        # Sensor priority
\end{Verbatim}

\paragraph{Native input files format}

The files \texttt{.blast}$<$i$>$ describe a collection of similarities
sorted by protein and by position on the sequence. One similarity S is
described per line.

The format of a line is:

\texttt{<$b^S$> <$e^S$> <$s^S$> <$v^S$> <$p^S$> <protein name> <$bp^S$> <$ep^S$>}

where:
\begin{itemize}
\item $b^S$ and $e^S$ are the begin and the end of the similarity S on the sequence,
\item $s^S$ is the score of the similarity S,
\item $v^S$ is the e-value given by BlastX and ignored by \EuGene,
\item $p^S$ is the phase: +1, +2, +3, -1, -2, -3,
\item $bp^S$ and $ep^S$are the begin and the end of the similarity S on the protein.
\end{itemize}

Here is an extract from \texttt{SYNO\_ARATH.fasta.blast0}:
\begin{Verbatim}[fontsize=\small]
2820 2861 36 3e-08 +3  sp_O07683_SYD_HALSA; 335 348
2972 3088 41 3e-08 +2  sp_O07683_SYD_HALSA; 359 397
3185 3298 113 3e-08 +2  sp_O07683_SYD_HALSA; 398 435
353 418 45 2e-13 +2  sp_O24822_SYD_HALVO; 13 34
1850 1915 67 2e-13 +2  sp_O24822_SYD_HALVO; 202 223
2775 2858 72 2e-13 +3  sp_O24822_SYD_HALVO; 318 345
3191 3280 104 2e-13 +2  sp_O24822_SYD_HALVO; 397 426
353 418 51 7e-12 +2  sp_O26328_SYD_METTH; 21 42
1271 1414 70 7e-12 +2  sp_O26328_SYD_METTH; 141 188
1850 1954 62 7e-12 +2  sp_O26328_SYD_METTH; 210 244
3191 3280 93 7e-12 +2  sp_O26328_SYD_METTH; 401 430
\end{Verbatim}
These files can be obtained directly from the output BlastX files by
parsing them with the \texttt{blast\_parser.pl} script.
The BlastX is launched with the command:
\begin{Verbatim}
  blastall -p blastx -d DATABASE_MULTIFASTA_PROTEIC_FILE -g F -F T -b
  500000 -v 500000 -e 1e-6 -i QUERY_GENOMIC_SEQUENCE_FASTA >
  TEMPORY_BLAST_RESULT_FILE
\end{Verbatim}
and the final \texttt{.blast}$<i>$ files are obtained with:
\begin{Verbatim}
  blast_parser.pl TEMPORY_BLAST_RESULT_FILE | sort -n -k 1,1 | sort -s
  -k 6,6 > QUERY_GENOMIC_SEQUENCE_FASTA.blast0
\end{Verbatim}

For more explanation, see the README file in the directory
\texttt{eugene/src/SensorPlugins/BlastX/GetData}.

\paragraph{Gff3 input file format}

The gff3 input mode is activated by setting the value \texttt{GFF3}
for the parameter \texttt{BlastX.format} in the parameter file.  The
plugin reads a collection of similarities from files named with the
sequence name and an extension .blast$<$i$>$.gff3
(\texttt{.blast0.gff3,...,.blast9.gff3}).

Each similarity is described by a GFF3 line as follows

\texttt{<sequence name> <program> <sofa feature> <$b^S$> <$e^S$> <$v^S$> <strand> <.> ID=...;Target=<protein name> <$bp^S$> <$ep^S$>;frame\_hit=<$p^S$>;score\_hit=<$s^S$>}

Gff3 attributes specifications:\\
Required attributes:\\
	\begin{itemize}
	\item ID
	\item Target
	\end{itemize}

Optional attributes:\\
	\begin{itemize}
	\item frame\_hit represent the blast frame. This information
          can be omitted, but if it's specified, eugene will check it
          is consistent with the similarity positions and reject the
          information otherwise.
	\item score\_hit :  normalized score.
	\end{itemize}

Here an extract of : seq14ac002535g4g5.tfa.blast0.gff3
\begin{Verbatim}[fontsize=\tiny]
seq14	blastx0	match	2361	4506	.	+	.	ID=blastx0:seq14.1;Dbxref=blastx0:sp_O84903_GALE_LACCA 
seq14	blastx0	match_part	2361	2483	2e-07	+	.	ID=blastx0:seq14.1.1;Dbxref=blastx0:sp_O84903_GALE_LACCA;Target=sp_O84903_GALE_LACCA 1 41 +;frame_hit=+3;score_hit=114;Parent=blastx0:seq14.1 
seq14	blastx0	match_part	2792	2857	2e-07	+	.	ID=blastx0:seq14.1.2;Dbxref=blastx0:sp_O84903_GALE_LACCA;Target=sp_O84903_GALE_LACCA 111 132 +;frame_hit=+2;score_hit=47;Parent=blastx0:seq14.1 
seq14	blastx0	match_part	4432	4506	2e-07	+	.	ID=blastx0:seq14.1.3;Dbxref=blastx0:sp_O84903_GALE_LACCA;Target=sp_O84903_GALE_LACCA 280 304 +;frame_hit=+1;score_hit=50;Parent=blastx0:seq14.1
\end{Verbatim}

\paragraph{Filtering input information}

Similarities with a length higher than 15,000 nucleodites are
rejected. A message ``Similarity of extreme length rejected'' is
printed to alert the user.

\paragraph{Integration of information}

The procedure consists first, in computing information at the
nucleotide level and second, in weigthing the graph used by \EuGene.

A/ Computing information at the nucleotide level

A-1/ Extracting information

Each similarity S is considered, one after the other.  A set of 3
variables is computed for nucleotide in position $i$. The variables
are:
\begin{itemize}
\item $s_i$ the score of the nucleotide at position i
\item $c_i$ the confidence in $s_i$,
\item $p_i$ the phase of $s_i$ : +1, +2, +3, -1, -2, -3 for exon and 0 for intron,
\end{itemize}
Let $l^S$ be the length of the similarity in nucleotide.
\[l^S = (ep^S -bp^S -1)*3\]

\underline{Valuation for exon position} \\

\begin{itemize}
\item from $i = b^S$ to  $i = e^S$
  \begin{itemize}
  \item $s_i = s^S / l^S$
  \item $c_i = c^S$
  \item $p_i = p^S$
  \end{itemize}
\end{itemize}

\underline{Valuation for intron position}\\

An intron is consider as more likely when there is sufficient gap between 2 HSP 
on the genomic sequence and the corresponding positions on the same 
protein are close together:
\begin{itemize}
\item on the protein side,  two HSP of the same protein appear on the same strand
  with a  maximum gap or overlap bteween the HSPs of \texttt{BlastX.blastxM*} amino acids.
\item on the genomic side, the length of the gap (in nucleotide) between 
the two HSPs is larger than \texttt{BlastX.minIn}.
\end{itemize}
Considering a gap S between 2 HSPs in these conditions, 
intron predictions are favored on a small region of
length \texttt{BlastX.minIn}/2 on both sides of the gap (inside the gap). 
\begin{itemize}
\item from $i = b^S -$ BlastX.minIn$/2$ to $i = b^S$ 
\item from $i = e^S$ to $i = e^S +$ BlastX.minIn$/2$ 
\item the following values are given at each position:
  \begin{itemize}
  \item $s_i = s^S / l^S$ 
  \item $c_i = c^S$ 
  \item $p_i = 0$
  \end{itemize}
\end{itemize}

If furthermore, the \texttt{activegaps} parameter is set for a level, then for all the bases
in the gap S which have not already been considered in the step above, the
intergenic, UTR and UTR introns states are also penalized. The penality used is the minimum
of the weights used for the left and right HSP around the gap S.

A-2/ Combining extracted information\\

When all the similarities have been handled, if a position has several
set of variables, the set with the highest confidence is kept. In
case of egal confidence, the set with the higher score is kept.

B/ Weighting the graph\\

For each $i$ with a set of variables:
\begin{itemize}
\item if $p_i$ codes for exon then $s_i .c_i$ is added to the content
  score of nucleotide i in the corresponding exon phase (a track
  between 0 and 5),
\item if $p_i$ codes for intron then $s_i .c_i$ is added to the intron
  score of nucleotide i (tracks 6 and 7),
\end{itemize}

Note: in fact, instead of rewarding the correct track (like described
here), all the tracks except the according one(s) are penalized, with
a penalty equal to $-|s_i .c_i|$.

\paragraph{Post analyse}

The correspondance between BlastX information and prediction is
analyzed if the \texttt{-B} flasg \index{CmdFlags}{[BlastX
  postprocessing activation] B} is provided or if the
\texttt{BlastX.PostProcess} parameter is set to \texttt{1}.

For each predicted CDS, from the start codon to the stop, the
percentage of nucleotides supported by a proteic similarity is
displayed.

\paragraph{Graph}

Grey horizontal lines are ploted on the exon tracks for only the 3
first levels to consider (dark grey for the first, grey for the
second, and light grey for the third).

