% Documentation of the PepSignal sensor

\subsubsection{\texttt{Sensor.PepSignal}}


\paragraph{Description}
This plugin injects possible translation starts as predicted by the Predotar
 program (http://genoplante-info.infobiogen.fr/predotar/) that looks for 
peptide adressing sequences after every occurrence of an ATG. 

The sensor is activated by setting the value \texttt{1} for the parameter
\texttt{Sensor.PepSignal.use} in the parameter file. The score for start
 prediction is rescaled by the parameters {\tt PepSignal.startP*} and
{\tt PepSignal.startB*} (see below).

Here is an example of PepSignal parameters definition.
\begin{Verbatim}[fontsize=\small]
PepSignal.startP*       0.9
PepSignal.startB*       0.1
Sensor.PepSignal.use    1       # Use PepSignal sensor	
Sensor.PepSignal        10      # Sensor priority
\end{Verbatim}


\paragraph{Native input files format}
The plugin reads 
the predictions (both in forward and reverse strand) of the program 
from a file whose name is derived 
from the sequence name by adding the \texttt{.psignal} suffix.

The file \texttt{.psignal} describes the predicted start sites
sorted by position on the sequence (as given by Predotar).
The format of a line is : \texttt{<position> <strand> <score> <comment>}

Here is an example used for test:
\begin{Verbatim}[fontsize=\small]
175 start_rev 0.024083 test
188 start 0.000151 test
195 start_rev 0.010081 test
261 start 0.001628 test
270 start 0.000026 test
[...]
\end{Verbatim}

This file can be obtained by launching the Predotar software
available at\\
\texttt{http://genoplante-info.infobiogen.fr/predotar/predotar.html}

\paragraph{Gff3 input file format}

The gff3 input mode is activated by setting the value \texttt{GFF3}
for the parameter \texttt{PepSignal.format} in the parameter file.
The plugin reads the predictions of the program from one file which
name is derived from the sequence name by adding the
\texttt{.psignal.gff3} extension.

Accepted features (third column):\\
\begin{itemize}
\item  SO:0000318 or start\_codon
\end{itemize}
If the feature used isn't one of these, the line will be rejected. The
expected coordinates must match with the first nucleotide of a
start codon (ATG).

Here an extract of \texttt{seq14ac002535g4g5.tfa.psignal.gff3}:
\begin{Verbatim}[fontsize=\tiny]
seq14	PepSignal	SO:0000318	175	175	0.024083	-	.	ID=SO:0000318:seq14.0;
seq14	PepSignal	SO:0000318	188	188	0.000151	+	.	ID=SO:0000318:seq14.1;
\end{Verbatim}

\paragraph{Filtering input information}

No filtering.


\paragraph{Integration of information}

The score (column 3) read $s$ is rescaled using the {\tt PepSignal.startP*} 
(\emph{P}) and {\tt PepSignal.startB*} (\emph{B}) as follows:

\[s' = e^{-P}.s^B\]

All predictions that use a predicted start receive a $\log(s')$
penalty while those that go through a predicted start while they
could have used it receive a $\log(1-s')$ penalty.


\paragraph{Post analyse}

No post analyse.


\paragraph{Graph}

Predicted starts are visible on exonix tracks as blue vertical lines
whose length indicates the site score.

