% Documentation of the ATGpr sensor

\subsubsection{\texttt{Sensor.ATGpr}}

\paragraph{Description}

This plugin injects possible translation starts as predicted by the
ATGpr program. This program is not available anymore and the plugin is 
no more supported and has been removed from recent versions of EuGene.

The plugin reads the prediction of the programs from two files whose
names are derived from the sequence name by adding the \texttt{.atgpr}
and \texttt{.atgprR} suffix (respectively prediction for the forward
and reverse strand).

The sensor is activated by setting the value 1 for the parameter
\texttt{Sensor.ATGpr.use} in the parameter file. The score for acceptor
and donor prediction is rescaled by the parameters {\tt ATGpr.startP*} and
{\tt ATGpr.startB*} (see below).

Here is an example of ATGpr parameters definition.
\begin{Verbatim}[fontsize=\small]
ATGpr.startP*      0.002
ATGpr.startB*      0.001
Sensor.ATGpr.use   1       # Use ATGpr sensor
Sensor.ATGpr       1       # Sensor priority
\end{Verbatim}

\paragraph{Input files format}

The files with \texttt{.atgpr} and \texttt{.atgprR} suffix are
obtained by running ATGpr which can be obtained at
\texttt{http://www.hri.co.jp/atgpr/} and using the
detailed output of the software.

Here is an extract from \texttt{SYNO\_ARATH.fasta.atgpr}:
\begin{Verbatim}[fontsize=\small]
1       0.26    2       AXXATGG 29      460     144     Yes     
2       0.12    3       tXXATGc 168     302     45      Yes     
3       0.04    1       GXXATGc 196     234     13      Yes     
4       0.04    3       GXXATGG 237     302     22      Yes     
5       0.04    2       AXXATGG 371     460     30      Yes     
6       0.04    2       tXXATGa 518     550     11      Yes     
7       0.04    2       AXXATGG 521     550     10      Yes     
8       0.04    3       tXXATGc 528     575     16      Yes     
9       0.04    3       GXXATGt 546     575     10      Yes     
10      0.04    1       GXXATGa 553     597     15      Yes     
11      0.04    2       cXXATGc 563     583     7       Yes     
12      0.04    2       GXXATGa 575     583     3       Yes     
13      0.04    3       cXXATGa 645     692     16      Yes     
14      0.04    1       GXXATGt 661     699     13      Yes     
15      0.04    2       GXXATGt 812     910     33      Yes     
[...]
\end{Verbatim}

\paragraph{Filtering input information}

No filtering.

\paragraph{Integration of information}

The integrated score for start prediction is read (column 2).  The
score read $s$ is rescaled using the {\tt ATGpr.startP*} (\emph{P}) and {\tt
ATGpr.startB*} (\emph{B}) as follows:

\[s' = e^{-P}*s^B\]

All predictions that use a predicted start receive a $\log(s')$
penalty while those that go through a predicted start while they
could have used it receive a $\log(1-s')$ penalty.


\paragraph{Post analyse}

No post analyse.

\paragraph{Graph}

Predicted starts are visible on exonix tracks as blue vertical lines
whose length indicates the site score.
