% Documentation of the MarkovProt sensor

\subsubsection{\texttt{Sensor.MarkovProt}}

\paragraph{Description}

This plugin injects coding/non coding likelihood as modeled by proteic
Markov models. These models are defined in a matrices file (located in
the directory specified by the \texttt{EuGene.PluginsDir} parameter)
whose name is indicated by the \texttt{MarkovProt.matname} parameter.
The order of the Markov model must be given in
\texttt{MarkovProt.maxorder} while the actual order to use is set by
\texttt{MarkovProt.order}.

The plugin is controlled by two further parameters:
\texttt{MarkovProt.minGC} and \texttt{MarkovProt.maxGC} which indicate
the GC scope of the matrices. If the GC\% of the sequence is out of
the scope, the plugin will give an equal null loglikelihood to all types
of regions.


The sensor is activated by either:
\begin{itemize}
\item the \texttt{-M} argument followed by the filename of the set of models.
  \index{CmdFlags}{[MarkovProt Markov matrix file] M}
\item the value 1 for the parameter \texttt{[Sensor.MarkovProt.use} in
  the parameter file.
\end{itemize}

Here is an example of MarkovProt parameters definition.
\begin{Verbatim}[fontsize=\small]
MarkovProt.matname      SwP41.noFragm.mininfo1.order2.bin
MarkovProt.minGC        0
MarkovProt.maxGC        100
MarkovProt.maxorder     2
MarkovProt.order        2
Sensor.MarkovProt.use   1    # Use MarkovProt sensor
Sensor.MarkovProt       1       # Sensor priority
\end{Verbatim}

\paragraph{Input files format}

No input files  needed beyond the markov matrix files.

\paragraph{Integration of information}

For coding tracks, assuming a uniform codon usage, the probability of
the coding tracks is decomposed as the product of choosing a codon and
then emitting the corresponding amino acid in the corresponding
phase. The logarithm of the probability is used for weighting.

For other tracks, a simple GC\% model is used to compute a background
probability. The logarithm of the probability is used for weighting.

\paragraph{Post analyse}

No post analyse.

\paragraph{Graph}

Same as in the MarkovIMM plugin.  


